Structured Versus Decorated Cospans

### 2020-11-11 cospans.tex
changed \ref to \cref everywhere

%{\chris Amend at the end}
%The outline for the paper is as follows: In Section \ref{DecCospansDoubleCat}, we construct the symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$, obtain the underlying symmetric monoidal bicategory $F \mathbf{Csp}$ via a result of Shulman, and decategorify this symmetric monoidal bicategory to obtain the symmetric monoidal category $F$Csp. In Section \ref{MapsDecCospansDoubleCat} we define maps between decorated cospan double categories which are double functors of an appropriate type. In Section \ref{some other section} we investigate when a symmetric lax monoidal pseudofunctor gives rise to a left adjoint via the Grothendieck construction, which naturally leads to another compositional framework known as `Structured cospans'. In Section \ref{EquivDoubleCats} we briefly review the framework Structured cospans and prove that the double categorical versions of decorated cospans and structured cospans are equivalent when both are appropriate. Finally, in Section \ref{Applications}, we present a few examples that can be realized with either framework and discuss the realized equivalences.

@@

%Let $F \maps \mathsf{A} \to \Cat$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \ref{DC} of the previous section, we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$. This symmetric monoidal double category has:
%\begin{enumerate}
%\item{objects as those of $\mathsf{A}$,}
%\item{vertical 1-morphisms as morphisms of $\mathsf{A}$,}
%\item{horizontal 1-cells as pairs:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$b$};
%\node (D) at (4,0) {$d \in F(c)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B);
%\end{tikzpicture}
%\]
%and}
%\item{2-morphisms as maps of cospans in $\mathsf{A}$
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$a$};
%\node (A') at (0,-0.5) {$a^\prime$};
%\node (B) at (1,0.5) {$c$};
%\node (C) at (2,0.5) {$b$};
%\node (C') at (2,-0.5) {$b^\prime$};
%\node (D) at (1,-0.5) {$c^\prime$};
%\node (E) at (3,0.5) {$d \in F(c)$};
%\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$$} (B)
%(C) edge node[above]{$$} (B)
%(A) edge node[left]{$f$} (A')
%(C) edge node[left]{$g$} (C')
%(A') edge node {$$} (D)
%(C') edge node {$$} (D)
%(B) edge node [left] {$h$} (D);
%\end{tikzpicture}
%\]
%together with a morphism $\tau \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
%\end{enumerate}

@@

where $\kappa$ is the isomorphism $$\kappa \maps H(c_1 +_b c_2) \xrightarrow{\sim} H(c_1) +_{H(b)} H(c_2)$$ which comes from the finite colimit preserving functor $H \maps \mathsf{A} \to \mathsf{A^\prime}$. The above diagram commutes by a similar argument to the one used in Theorem \ref{Equiv}. Similarly, for every object $c \in \mathsf{A}$, we have a unit comparison constraint $$\mathbb{H}_U \maps U_{\mathbb{H}(c)} \to \mathbb{H}(U_c)$$ given by the globular 2-isomorphism:

@@

The diagrams involving the morphisms of decorations are similar to those in Theorem \ref{DC}. This shows that $\mathbb{H}=(H,E,\theta)$ is a double functor. Next we show that this double functor is symmetric monoidal. First, that the object component $\mathbb{H}_0=H$ is symmetric monoidal is clear as $H \maps \mathsf{A} \to \mathsf{A}^\prime$ preserves finite colimits. As for the arrow component $\mathbb{H}_1$, given two horizontal 1-cells $M_1$ and $M_2$ in $F\mathbb{C}\textnormal{sp}$:

@@

\section{Applications}\label{Applications}
\begin{comment}
In this section we present several examples each of which may be realized in the context of decorated cospans or in the context of structured cospans. The first example regarding graphs was mentioned in the introduction and used as a reoccurring theme throughout the paper. The next three examples which take on more of an applied flavor, consists of electrical circuits, Markov processes and Petri nets. Each of these has been studied extensively by Baez, Fong, Master and Pollard by way of `black-boxing' \cite{BCR,BF,BFP,BM,BP}. Black-boxing is a way of interpreting the behavior of an open system, that is, a system with prescribed inputs and outputs such as the terminals of an electrical circuit, by observing the activity at the inputs and the outputs. The semantics of the activity at an open system's inputs and outputs is typically described in a category such as $\mathsf{LinRel}$ of finite dimensonal vector spaces and linear relations. Thus, in each case, black-boxing results in functors such as: $$\blacksquare_1 \maps \mathsf{Circ} \to \mathsf{LinRel}$$ $$\blacksquare_2 \maps \mathsf{Mark} \to \mathsf{LinRel}$$ $$\blacksquare_3 \maps \mathsf{Petri} \to \mathsf{LinRel}.$$ Each of these black-boxing functors also possess other convenient properties such as being symmetric monoidal. The first two of these were first done using Fong's theory of decorated cospans and then extended by the first two authors using structured cospans. The last two of these were also extended by being realized as double functors between double categories in recent works \cite{BC,BM}.
\subsection{Graphs}
As a first example that was also mentioned in the introduction, let $L \maps \mathsf{Set} \to \mathsf{Graph}$ be the functor that assigns to a set $N$ the \emph{discrete graph} on $N$ which is the edgeless graph $L(N)$ with no edges and $N$ as its set of vertices. Both $\mathsf{Set}$ and $\mathsf{Graph}$ are cocartesian monoidal and the functor $L \maps \mathsf{Set} \to \mathsf{Graph}$ is left adjoint to the forgetful functor $R \maps \mathsf{Graph} \to \mathsf{Set}$ which assigns to a graph $G$ its underlying set of vertices $U(G)$. Using structured cospans and appealing to Theorem \ref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{cospans of graphs, or, \emph{open} graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(M)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
as horizontal 1-cells, where $L(N)$ and $L(M)$ are discrete graphs on the sets $N$ and $M$, respectively, $G$ is a graph and $I$ and $O$ are graph morphisms, and}
\item{maps of cospans of graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(M_1)$};
\node (A') at (0,-1) {$L(N_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
as 2-morphisms, where $L(f)$ and $L(g)$ are maps of discrete graphs induced by the underlying functions $f$ and $g$, respectively, and $\alpha \maps G_1 \to G_2$ is a graph morphism.
}
\end{enumerate}

We can obtain a similar symmetric monoidal double category using decorated cospans. Let $F \maps \mathsf{Set} \to \mathsf{Cat}$ be the symmetric lax monoidal pseudofunctor that assigns to a set $N$ the \emph{category} of all graph structures whose underlying set of vertices is $N$. Using Theorem \ref{DC}, we then obtain a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{horizontal 1-cells as pairs:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$N$};
\node (B) at (1,0) {$P$};
\node (C) at (2,0) {$M$};
\node (D) at (3.25,0) {$G \in F(P)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
which can also be thought of as open graphs, and}
\item{2-morphisms as maps of cospans of sets
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$N_1$};
\node (A') at (0,-1) {$N_2$};
\node (C') at (2,-1) {$M_2$};
\node (B) at (1,0) {$P_1$};
\node (C) at (2,0) {$M_1$};
\node (D) at (1,-1) {$P_2$};
\node (E) at (3,0) {$G_1 \in F(P_1)$};
\node (F) at (3,-1) {$G_2 \in F(P_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(C') edge node [above] {$o_2$} (D)
(A') edge node [above] {$i_2$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a graph morphism $\tau \maps F(h)(G_1) \to G_2$ in $F(P_2)$.}
\end{enumerate}
We thus have two symmetric monoidal double categories: $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ obtained from structured cospans and $F\mathbb{C}\textnormal{sp}$ obtained from decorated cospans. Both of these double categories have $\mathsf{Set}$ as their categories of objects, open graphs as horizontal 1-cells and maps of open graphs as 2-morphisms, and by Theorem \ref{Equiv}, we have an equivalence of symmetric monoidal double categories $$_L \mathbb{C}\textnormal{sp}(\mathsf{Graph}) \sim F\mathbb{C}\textnormal{sp}.$$

@@

Thus the pseudofunctor $F$ is symmetric lax monoidal and so by Theorem \ref{DC} we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

These two double categories $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph}_k)$ and $F\mathbb{C}\textnormal{sp}$ are equivalent by Theorem \ref{Equiv}.

@@

This is done using a symmetric lax monoidal functor $F \maps \mathsf{FinSet} \to \mathsf{Set}$ which assigns to each finite set $N$ the (large) set of all Markov processes on $N$ as defined above. Viewing this functor $F$ as now a symmetric lax monoidal pseudofunctor $F \maps \mathsf{FinSet} \to \Cat$ that assigns to a finite set $N$ the \emph{category} $F(N)$ of all Markov processes on $N$, we then get by Theorem \ref{DC} a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

A symmetric monoidal double category of open Markov processes can also be obtained using structured cospans by defining a functor $L \maps \mathsf{FinSet} \to \mathsf{Mark}$ which assigns to a finite set $N$ the discrete Markov process $L(N)$ with no edges and to a function $f \maps N \to N^\prime$ the induced map of discrete Markov processes. Both categories $\mathsf{FinSet}$ and $\mathsf{Mark}$ have finite colimits and the functor $L$ is left adjoint to the forgetful functor $R \maps \mathsf{Mark} \to \mathsf{FinSet}$ which maps a Markov process to its underlying finite set of states. By Theorem \ref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Mark})$ which has:

@@

The two double categories $F\mathbb{C}\textnormal{sp}$ and $_L \mathbb{C} \textnormal{sp}(\mathsf{Mark})$ are equivalent by Theorem \ref{Equiv}.

@@

Each set of species $S$ gives rise to a discrete Petri net $L(S)$ with $S$ as its set of species and no transitions. Master and the first author show the existence of a left adjoint $L \maps \mathsf{Set} \to \mathsf{Petri}$ where $\mathsf{Petri}$ is the category whose objects are Petri nets and whose `morphisms are morphisms of Petri nets'. They also show that $\mathsf{Petri}$ has finite colimits and thus using Theorem \ref{SC} obtain a symmetric monoidal double category $\mathbb{O}\mathbf{pen}(\mathrm{Petri})$ of open Petri nets which has:

@@

We can also obtain a similar double category using decorated cospans: define a pseudofunctor $F \maps \mathsf{Set} \to \Cat$ where given a set $s$, $F(s)$ is the category of all Petri net structures with $s$ as its set of species. This pseudofunctor $F$ is symmetric lax monoidal as both $(\mathsf{Set},+,\emptyset)$ and $(\Cat,\times,1)$ are symmetric monoidal and given Petri nets $P \in F(s)$ and $P^\prime \in F(s^\prime)$, we can place them side by side and consider them together as a single Petri net $P+P^\prime \in F(s+s^\prime)$ with set of species $s+s^\prime$, and thus we have natural transformations $\phi_{s,s^\prime} \maps F(s) \times F(s^\prime) \to F(s+s^\prime)$ for any two sets $s$ and $s^\prime$. The other structure morphism between monoidal units $\phi \maps 1_{\mathsf{Petri}} \to F(\emptyset)$ is defined by the unique morphism from the empty Petri net with the empty set forits set of species to the only possible Petri net on the empty set, which is also the empty Petri net. All of the diagrams that are required to commute are straightforward. Using Theorem \ref{DC}, we obtain a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

Thus we have a symmetric monoidal double category $\mathbb{O}\mathbf{pen}(\mathrm{Petri})$ of open Petri nets obtained from structured cospans and a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of open Petri nets obtain from decorated cospans, and of course, we have an equivalence $\mathbb{O}\mathbf{pen}(\mathrm{Petri}) \sim F\mathbb{C}\textnormal{sp}$ of symmetric monoidal double categories by Theorem \ref{Equiv}.

>>>>

%{\chris Amend at the end}
%The outline for the paper is as follows: In Section \cref{DecCospansDoubleCat}, we construct the symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$, obtain the underlying symmetric monoidal bicategory $F \mathbf{Csp}$ via a result of Shulman, and decategorify this symmetric monoidal bicategory to obtain the symmetric monoidal category $F$Csp. In Section \cref{MapsDecCospansDoubleCat} we define maps between decorated cospan double categories which are double functors of an appropriate type. In Section \cref{some other section} we investigate when a symmetric lax monoidal pseudofunctor gives rise to a left adjoint via the Grothendieck construction, which naturally leads to another compositional framework known as `Structured cospans'. In Section \cref{EquivDoubleCats} we briefly review the framework Structured cospans and prove that the double categorical versions of decorated cospans and structured cospans are equivalent when both are appropriate. Finally, in Section \cref{Applications}, we present a few examples that can be realized with either framework and discuss the realized equivalences.

@@

%Let $F \maps \mathsf{A} \to \Cat$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \cref{DC} of the previous section, we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$. This symmetric monoidal double category has:
%\begin{enumerate}
%\item{objects as those of $\mathsf{A}$,}
%\item{vertical 1-morphisms as morphisms of $\mathsf{A}$,}
%\item{horizontal 1-cells as pairs:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$b$};
%\node (D) at (4,0) {$d \in F(c)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B);
%\end{tikzpicture}
%\]
%and}
%\item{2-morphisms as maps of cospans in $\mathsf{A}$
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$a$};
%\node (A') at (0,-0.5) {$a^\prime$};
%\node (B) at (1,0.5) {$c$};
%\node (C) at (2,0.5) {$b$};
%\node (C') at (2,-0.5) {$b^\prime$};
%\node (D) at (1,-0.5) {$c^\prime$};
%\node (E) at (3,0.5) {$d \in F(c)$};
%\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$$} (B)
%(C) edge node[above]{$$} (B)
%(A) edge node[left]{$f$} (A')
%(C) edge node[left]{$g$} (C')
%(A') edge node {$$} (D)
%(C') edge node {$$} (D)
%(B) edge node [left] {$h$} (D);
%\end{tikzpicture}
%\]
%together with a morphism $\tau \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
%\end{enumerate}

@@

where $\kappa$ is the isomorphism $$\kappa \maps H(c_1 +_b c_2) \xrightarrow{\sim} H(c_1) +_{H(b)} H(c_2)$$ which comes from the finite colimit preserving functor $H \maps \mathsf{A} \to \mathsf{A^\prime}$. The above diagram commutes by a similar argument to the one used in Theorem \cref{Equiv}. Similarly, for every object $c \in \mathsf{A}$, we have a unit comparison constraint $$\mathbb{H}_U \maps U_{\mathbb{H}(c)} \to \mathbb{H}(U_c)$$ given by the globular 2-isomorphism:

@@

The diagrams involving the morphisms of decorations are similar to those in Theorem \cref{DC}. This shows that $\mathbb{H}=(H,E,\theta)$ is a double functor. Next we show that this double functor is symmetric monoidal. First, that the object component $\mathbb{H}_0=H$ is symmetric monoidal is clear as $H \maps \mathsf{A} \to \mathsf{A}^\prime$ preserves finite colimits. As for the arrow component $\mathbb{H}_1$, given two horizontal 1-cells $M_1$ and $M_2$ in $F\mathbb{C}\textnormal{sp}$:

@@

\section{Applications}\label{Applications}
\begin{comment}
In this section we present several examples each of which may be realized in the context of decorated cospans or in the context of structured cospans. The first example regarding graphs was mentioned in the introduction and used as a reoccurring theme throughout the paper. The next three examples which take on more of an applied flavor, consists of electrical circuits, Markov processes and Petri nets. Each of these has been studied extensively by Baez, Fong, Master and Pollard by way of `black-boxing' \cite{BCR,BF,BFP,BM,BP}. Black-boxing is a way of interpreting the behavior of an open system, that is, a system with prescribed inputs and outputs such as the terminals of an electrical circuit, by observing the activity at the inputs and the outputs. The semantics of the activity at an open system's inputs and outputs is typically described in a category such as $\mathsf{LinRel}$ of finite dimensonal vector spaces and linear relations. Thus, in each case, black-boxing results in functors such as: $$\blacksquare_1 \maps \mathsf{Circ} \to \mathsf{LinRel}$$ $$\blacksquare_2 \maps \mathsf{Mark} \to \mathsf{LinRel}$$ $$\blacksquare_3 \maps \mathsf{Petri} \to \mathsf{LinRel}.$$ Each of these black-boxing functors also possess other convenient properties such as being symmetric monoidal. The first two of these were first done using Fong's theory of decorated cospans and then extended by the first two authors using structured cospans. The last two of these were also extended by being realized as double functors between double categories in recent works \cite{BC,BM}.
\subsection{Graphs}
As a first example that was also mentioned in the introduction, let $L \maps \mathsf{Set} \to \mathsf{Graph}$ be the functor that assigns to a set $N$ the \emph{discrete graph} on $N$ which is the edgeless graph $L(N)$ with no edges and $N$ as its set of vertices. Both $\mathsf{Set}$ and $\mathsf{Graph}$ are cocartesian monoidal and the functor $L \maps \mathsf{Set} \to \mathsf{Graph}$ is left adjoint to the forgetful functor $R \maps \mathsf{Graph} \to \mathsf{Set}$ which assigns to a graph $G$ its underlying set of vertices $U(G)$. Using structured cospans and appealing to Theorem \cref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{cospans of graphs, or, \emph{open} graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(M)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
as horizontal 1-cells, where $L(N)$ and $L(M)$ are discrete graphs on the sets $N$ and $M$, respectively, $G$ is a graph and $I$ and $O$ are graph morphisms, and}
\item{maps of cospans of graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(M_1)$};
\node (A') at (0,-1) {$L(N_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
as 2-morphisms, where $L(f)$ and $L(g)$ are maps of discrete graphs induced by the underlying functions $f$ and $g$, respectively, and $\alpha \maps G_1 \to G_2$ is a graph morphism.
}
\end{enumerate}

We can obtain a similar symmetric monoidal double category using decorated cospans. Let $F \maps \mathsf{Set} \to \mathsf{Cat}$ be the symmetric lax monoidal pseudofunctor that assigns to a set $N$ the \emph{category} of all graph structures whose underlying set of vertices is $N$. Using Theorem \cref{DC}, we then obtain a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{horizontal 1-cells as pairs:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$N$};
\node (B) at (1,0) {$P$};
\node (C) at (2,0) {$M$};
\node (D) at (3.25,0) {$G \in F(P)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
which can also be thought of as open graphs, and}
\item{2-morphisms as maps of cospans of sets
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$N_1$};
\node (A') at (0,-1) {$N_2$};
\node (C') at (2,-1) {$M_2$};
\node (B) at (1,0) {$P_1$};
\node (C) at (2,0) {$M_1$};
\node (D) at (1,-1) {$P_2$};
\node (E) at (3,0) {$G_1 \in F(P_1)$};
\node (F) at (3,-1) {$G_2 \in F(P_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(C') edge node [above] {$o_2$} (D)
(A') edge node [above] {$i_2$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a graph morphism $\tau \maps F(h)(G_1) \to G_2$ in $F(P_2)$.}
\end{enumerate}
We thus have two symmetric monoidal double categories: $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ obtained from structured cospans and $F\mathbb{C}\textnormal{sp}$ obtained from decorated cospans. Both of these double categories have $\mathsf{Set}$ as their categories of objects, open graphs as horizontal 1-cells and maps of open graphs as 2-morphisms, and by Theorem \cref{Equiv}, we have an equivalence of symmetric monoidal double categories $$_L \mathbb{C}\textnormal{sp}(\mathsf{Graph}) \sim F\mathbb{C}\textnormal{sp}.$$

@@

Thus the pseudofunctor $F$ is symmetric lax monoidal and so by Theorem \cref{DC} we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

These two double categories $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph}_k)$ and $F\mathbb{C}\textnormal{sp}$ are equivalent by Theorem \cref{Equiv}.

@@

This is done using a symmetric lax monoidal functor $F \maps \mathsf{FinSet} \to \mathsf{Set}$ which assigns to each finite set $N$ the (large) set of all Markov processes on $N$ as defined above. Viewing this functor $F$ as now a symmetric lax monoidal pseudofunctor $F \maps \mathsf{FinSet} \to \Cat$ that assigns to a finite set $N$ the \emph{category} $F(N)$ of all Markov processes on $N$, we then get by Theorem \cref{DC} a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

A symmetric monoidal double category of open Markov processes can also be obtained using structured cospans by defining a functor $L \maps \mathsf{FinSet} \to \mathsf{Mark}$ which assigns to a finite set $N$ the discrete Markov process $L(N)$ with no edges and to a function $f \maps N \to N^\prime$ the induced map of discrete Markov processes. Both categories $\mathsf{FinSet}$ and $\mathsf{Mark}$ have finite colimits and the functor $L$ is left adjoint to the forgetful functor $R \maps \mathsf{Mark} \to \mathsf{FinSet}$ which maps a Markov process to its underlying finite set of states. By Theorem \cref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Mark})$ which has:

@@

The two double categories $F\mathbb{C}\textnormal{sp}$ and $_L \mathbb{C} \textnormal{sp}(\mathsf{Mark})$ are equivalent by Theorem \cref{Equiv}.

@@

Each set of species $S$ gives rise to a discrete Petri net $L(S)$ with $S$ as its set of species and no transitions. Master and the first author show the existence of a left adjoint $L \maps \mathsf{Set} \to \mathsf{Petri}$ where $\mathsf{Petri}$ is the category whose objects are Petri nets and whose `morphisms are morphisms of Petri nets'. They also show that $\mathsf{Petri}$ has finite colimits and thus using Theorem \cref{SC} obtain a symmetric monoidal double category $\mathbb{O}\mathbf{pen}(\mathrm{Petri})$ of open Petri nets which has:

@@

We can also obtain a similar double category using decorated cospans: define a pseudofunctor $F \maps \mathsf{Set} \to \Cat$ where given a set $s$, $F(s)$ is the category of all Petri net structures with $s$ as its set of species. This pseudofunctor $F$ is symmetric lax monoidal as both $(\mathsf{Set},+,\emptyset)$ and $(\Cat,\times,1)$ are symmetric monoidal and given Petri nets $P \in F(s)$ and $P^\prime \in F(s^\prime)$, we can place them side by side and consider them together as a single Petri net $P+P^\prime \in F(s+s^\prime)$ with set of species $s+s^\prime$, and thus we have natural transformations $\phi_{s,s^\prime} \maps F(s) \times F(s^\prime) \to F(s+s^\prime)$ for any two sets $s$ and $s^\prime$. The other structure morphism between monoidal units $\phi \maps 1_{\mathsf{Petri}} \to F(\emptyset)$ is defined by the unique morphism from the empty Petri net with the empty set forits set of species to the only possible Petri net on the empty set, which is also the empty Petri net. All of the diagrams that are required to commute are straightforward. Using Theorem \cref{DC}, we obtain a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:

@@

Thus we have a symmetric monoidal double category $\mathbb{O}\mathbf{pen}(\mathrm{Petri})$ of open Petri nets obtained from structured cospans and a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of open Petri nets obtain from decorated cospans, and of course, we have an equivalence $\mathbb{O}\mathbf{pen}(\mathrm{Petri}) \sim F\mathbb{C}\textnormal{sp}$ of symmetric monoidal double categories by Theorem \cref{Equiv}.
